6.890 Spring 2019 Project Proposal \\ LHYRA: Learned HYbrid Recursive Algorithms

### 2019-05-16 proposal/proposal.tex

Practically, what usually happens is that an expert programmer writes a number of these algorithms, benchmarks them against each other, and chooses the one which appears to perform the best. A clear example of this is in the C++ STL, in which the standard algorithm is Introsort, which calls Quicksort until hitting a certain recursion depth, where it switches to Heapsort. We observe that it would be better that these recursive structures would adapt to their distributions and objective functions, so that not only could they lessen the workload for the expert, but also better explore the potential synergies of the possible algorithms.

@@

Regardless of the results, our expansive benchmarking will allow us to determine which data structures and applications are most improved by our use of multiple solvers (ignoring the time spend in the decision making process). This information could lead to future advancements both in application specific tasks (where the behavior of our solver selector could be manually replicated) and in generalized cases, as more efficient solver selectors / feature extractors are found.

@@

If one were to think of the most naive possible way to do this, you could imagine randomly sampling the bag of solvers (including their hyperparameters) and the training data to build a table of (feature set, solver) $\rightarrow$ cost. One could then perform a variant of nearest neighbor on that table given an input $x$ to find the solver with the lowest cost.

However, this approach is rife with complication. First, it is difficult using this framework to identify relationships between solvers (for example, if one solver is very good at solving its own subproblems). It would even be difficult to individually measure solver performance, as a fast solver that calls big subproblems might actually be less efficient than a slow solver that calls small subproblems. Even if one tries to get around this issue by averaging subproblem approaches, one still runs into the above issue of failing to model more complex relationships between solvers and the problems they best perform on.

@@

\begin{itemize}
	\item \textbf{B-Trees.} If possible, we would like to re-implement Kraska’s B-Trees in LHYRA to demonstrate how LHYRA is a true generalization. As LHYRA is partially modelled after RMI, we expect this to be a fairly straightforward adaptation.
	\item \textbf{Sorting.} We’d like to show how LHYRA can adapt to different distributions in choosing the optimal sorting method. We’d also like to show that LHYRA can also adapt to quickly sort approximately, and utilize processor parallelism where available.
	\item \textbf{Hierarchical Pathfinding.} Traditional pathfinding algorithms waste substantial amounts of computation and have difficulty adjusting to dynamic, online environments. The method of choice for hierarchical pathfinding is likely highly dependent on the exact optimization criteria, graph connectivity, and dynamism of the environment. This is a place where LHYRA may be invaluable, and also has clear real-world applications.
	\item \textbf{Matrix Multiplication.} LHYRA should be able to adaptively select naive matrix multiplication, naive divide and conquer, Strassen’s algorithm, or Coppersmith–Winograd. This would be particularly useful for machine learning.
	Nearest Neighbor: Given high-dimensional data, find the nearest neighbor. Examples of methods for this could be Locality sensitive hashing, Greedy search in proximity neighborhoods, and naive search.

@@

This is a project with a lot of breadth, flexibility, and engineering effort, which we believe warrants a three person group. We intend to develop LHYRA in Python for the sake of implementation speed, as we can still demonstrate performance improvements over naive Python implementations.

>>>>

Practically, what usually happens is that an expert programmer writes a number of these algorithms, benchmarks them against each other, and chooses the one which appears to perform the best. A clear example of this is in the C++ STL, in which the standard algorithm is Introsort, which calls Quicksort until hitting a certain recursion depth, where it switches to Heapsort. We observe that it would be better if these recursive structures adapted to their distributions and objective functions, so that not only could they lessen the workload for the expert, but also better explore potential algorithmic `synergies'.

@@

Regardless of the results, our expansive benchmarking will allow us to determine which data structures and applications are most improved by our use of multiple solvers (ignoring the time spend in the decision making process). This information could lead to future advancements both in application specific tasks, where the behavior of our solver selector could be manually replicated, and in generalized cases, as more efficient solver selectors / feature extractors are found.

@@

The most naive possible way to do this would involve randomly sampling the bag of solvers (including their hyperparameters) and the training data to build a table of (feature set, solver) $\rightarrow$ cost. We could then perform a variant of nearest neighbor on that table given an input $x$ to find the solver with the lowest cost.

However, this approach is rife with complication. First, it is difficult to individually measure solver performance, as a fast solver that calls big subproblems might actually be less efficient than a slow solver that calls small subproblems. Even if we try to get around this issue by averaging subproblem approaches, we would fail to model more complex relationships between solvers (for example, if one solver is particularly good at solving its own subproblems).

@@

\begin{itemize}
	\item \textbf{B-Trees.} If possible, we would like to re-implement Kraska’s B-Trees in LHYRA to demonstrate how LHYRA is a true generalization. As LHYRA is partially modelled after RMI, we expect this to be a fairly straightforward adaptation. Additionally, LHYRA should make it easier to experiment with additional CDF predictors.
	\item \textbf{Sorting.} We’'d like to show how LHYRA can adapt to different distributions in choosing the optimal sorting method. We’d also like to show that LHYRA can also adapt to quickly sort approximately, and utilize processor parallelism where available.
	\item \textbf{Hierarchical Pathfinding.} Traditional pathfinding algorithms waste substantial amounts of computation and have difficulty adjusting to dynamic, online environments. The method of choice for hierarchical pathfinding is likely highly dependent on the exact optimization criteria, graph connectivity, and dynamism of the environment. This is a place where LHYRA may be invaluable, and also has clear real-world applications.
	\item \textbf{Matrix Multiplication.} LHYRA should be able to adaptively select naive matrix multiplication, naive divide and conquer, Strassen’s algorithm, or Coppersmith-–Winograd. This would be particularly useful for machine learning.
	\item \textbf{Nearest Neighbor.} Given high-dimensional data, find the nearest neighbor. Examples of methods for this could be Locality sensitive hashing, Greedy search in proximity neighborhoods, and naive (linear) search.

@@

This is a project with a lot of breadth, flexibility, and engineering effort, which we believe warrants a three person group. We intend to develop LHYRA in Python for the sake of implementation speed, as we can still demonstrate performance improvements over naive Python implementations. Additionally, if we have sufficiently promising results from the Python implementation, we may also port several of the generated solutions to C/C++ to see how LHYRA compares to the state of the art.
